How to Not write Virtual Machines for Dynamic Languages

### 2007-05-27 talk/dyla2007/dyla.tex
removing some XXXs

In section XXX we will explore the way VMs are typically implemented
in C and some of the problems of this approach. In section XXX we will describe
our proposed meta-programming approach. XXX

\section{XXX XXX}

@@

As a language becomes more popular, yet another reason for different
implementations emerges: People want to use it together with Java and C\#
libraries and tools. Although it is in principle possible to do that by
bridging the C-VM with the respective VMs of these languages (XXX examples)
this is an incomplete solution since it prevents many use-cases.
Therefore in the end implementations of the dynamic language on top of the
object-oriented VMs are created to properly integrate with the complete
environment. Implementing a language on top of an OO VM is in many ways easier
than implementing it in C.

@@

Let's take a look at the advantages that are usually cited for basing a
language implementation of a dynamic language on a standard object oriented
virtual machine, for example the JVM or the CLR. The central theme of the
benefits  of OO VMs is the ability to implement certain hard things only once
and share the benefits between all language implementations on top of the OO VM.

@@

\emph{Better tools:} XXX

@@

\emph{Better performance:} So far it seems that performance of highly dynamic
languages is not actually significantly improved on OO VMs. Jython
\footnote{Python on the Java VM} is around 5
times slower than CPython, for IronPython\footnote{Python on .NET, which
gives up on some features to improve performance}
the figures vary but it is mostly
within the same order of magnitude as CPython. The most important reason for
this is that the VM's JIT compilers are not prepared to deal with the highly
dynamic behaviour and the complex dispatch semantics of dynamic languages. (XXX
expand)

@@

\section{Meta-Programming Is Good}

>>>>

In section \ref{sect:approaches} we will explore the way VMs are typically
implemented in C and on top of OO VMs and some of the problems of these
approaches. In \ref{section sect:metaprogramming} we will describe our proposed
meta-programming approach.

\section{Approaches to Dynamic Language Implementation}
\label{sect:appraches}

@@

As a language becomes more popular, yet another reason for different
implementations emerges: People want to use it together with Java and C\#
libraries and tools. Although it is in principle possible to do that by
bridging the C-VM with the respective VMs of these languages (XXX examples)
this is an incomplete solution since it prevents many use-cases.
Therefore in the end implementations of the dynamic language on top of the
object-oriented VMs are created to properly integrate with the complete
environment.

@@

XXX mention that OO languages don't have good metaprogramming support

Implementing a language on top of an OO VM is in many ways easier than
implementing it in C. Let's take a look at the advantages that are usually
cited for basing a
language implementation of a dynamic language on a standard object oriented
virtual machine, for example the JVM or the CLR. The central theme of the
benefits  of OO VMs is the ability to implement certain hard things only once
and share the benefits between all language implementations on top of the OO VM.

@@

\emph{Better tools:} Better IDEs, debuggers and profilers.

@@

\emph{Better performance:} So far it seems that performance of highly dynamic
languages is not actually significantly improved on OO VMs. Jython
\footnote{Python on the Java VM} is around 5
times slower than CPython, for IronPython\footnote{Python on .NET, which
gives up on some features to improve performance}
the figures vary but it is mostly
within the same order of magnitude as CPython. The most important reason for
this is that the VM's JIT compilers are optimized for specific usage patterns
that are common in the main language of the OO VM. To get good speeds the
language implementers would have to carefully produce code that matches these
usage patterns, which is not a simple task.

@@

\section{Meta-Programming Is Good}
\label{sect:metaprogramming}
